IML Term project paper

### 2022-12-23 paper.rmd
Update paper.rmd

```{r}

@@

## Description of considered machine learning approaches
In order to use a model that describes the process that creates the data the best, we attempted to use several different kinds of models to assess the data. We test these models on the original dataset or on a suitably modified version of it. After defining each of these models, we would assess which of the models we used performed best and use said model in our predictions.

@@

In tree-based methods, the data points are divided into smaller and smaller fragments based on some set of rules. A new split happens at each node, so when the size of the tree grows, it becomes more and more fitted to the data. When fitting classification trees, the process is often first started with a very large tree, and then the unnecessary branches are removed in the process of pruning, until the complexity of the tree is reduces to the optimal level. Constructing the trees, pruning, and evaluating the results can be done in a variety of different ways.

@@

#### Random Forest
One often used tree-based classification method is called Random Forest. It is based on the process described above, but only a random subset of the features is considered at each split of the tree. This is repeated multiple times and in the end, the obtained forest is evaluated and the best splits at each point are found. For Random Forest classifier, we used the R library *randomForest*.

We tried the model with both the full dataset and the dataset with 36 variables, of which the model with the fewer variables performed better. We used that model as our Random Forest classifier.

#### GBM
Another tree-based model that we used was GBM (gradient boosting machine). It starts by building a set of shallow trees, but with each iteration, it learns more about the data and improves each previously built tree. By finally combining the trees, it can produce very good models for all types of data sets. It's advantage is that it doesn't necessarily require any pre-processing of the data, although the model can be improved by this. For GBM classifier, we used the R library *GBM*.

@@

The results for the two-class models in terms of loo-cross-validated accuracy and perplexity are:
\begin{table}
\begin{tabular}{|r|ccc|}
\hline
Model type & Loo-CV accuracy & Loo-CV perplexity&Binary accuracy on test data & Binary accuracy on npf_test\\
\hline
Dummy model & 0 & 2.01&\\
Logistic regression & 0.8728448 & 1.338286&0.8445596\\
Naive Bayes & 0.8125 & 166.084 &\\
LDA & 0.8922414 & 1.32923 &\\
QDA & 0.8534483 & 6.34633 &\\
SVM & 0.8793103 & -&0.8435233\\
Lasso & 0.8965517 & 1.29399&0.8746114\\
Random Forest & 0.89 & 1.35\\
GBM &	0.8943966 & 1.324385 & 0.8746114 \\
\hline
\end{tabular}
\end{table}

@@

\begin{table}

\begin{tabular}{|r|c|}

\hline

Explanatory variable & Estimate of coefficient\\

\hline

Intercept&16.16886\\

CO2168.std&0.02065910\\

CO242.mean&-0.04024650\\

CO2504.mean&-0.02066992\\

H2O42.mean&-0.2627714\\

H2O84.std&-2.407117e-01\\

NO168.std&2.281875\\

NOx672.std&-0.06064781\\

O3672.mean&0.09233133\\

O384.std&0.03595632\\

Pamb0.std&0.2960355\\

PTG.mean&-15.19914\\

RGlob.mean&0.001578964\\

RHIRGA168.std&0.01765280\\

RHIRGA336.mean&-0.02661884\\

RHIRGA672.mean&-0.03528919\\

RHIRGA672.std&0.06932797\\

SO2168.std&0.7092859\\

SWS.mean&0.01112000\\

T168.std&0.4947142\\

CS.mean&-938.8368\\

CS.std&580.6009\\

\hline

\end{tabular}

\end{table}

@@

## Delivarables

**For the deliverables, we give a grade x.**

>>>>

```{r}0.8455959

@@

## Description of considered machine learning approaches

In order to use a model that describes the process that creates the data the best, we attempted to use several different kinds of models to assess the data. We test these models on the original dataset or on a suitably modified version of it. After defining each of these models, we would assess which of the models we used performed best and use said model in our predictions.

@@

In tree-based methods, the data points are divided into smaller and smaller fragments based on some set of rules. A new split happens at each node, so when the size of the tree grows, it becomes more and more fitted to the data. When fitting classification trees, the process is often first started with a very large tree, and then the unnecessary branches are removed in the process of pruning, until the complexity of the tree is reduces to the optimal level. Constructing the trees, pruning, and evaluating the results can be done in a variety of different ways.[@book, p. 328-361]

@@

#### Random Forest

One often used tree-based classification method is called Random Forest. It is based on the process described above, but only a random subset of the features is considered at each split of the tree. This is repeated multiple times and in the end, the obtained forest is evaluated and the best splits at each point are found. For Random Forest classifier, we used the R library *randomForest*.

We tried the model with different parameters and with both the full dataset and the dataset with 36 variables, of which the model with all the variables performed a little better. We used that model as our Random Forest classifier in the comparison.

#### GBM

Another tree-based model that we used was GBM (gradient boosting machine). It starts by building a set of shallow trees, but with each iteration, it learns more about the data and improves each previously built tree. By finally combining the trees, it can produce very good models for all types of data sets. It's advantage is that it doesn't necessarily require any pre-processing of the data, although the model can be improved by this. For GBM classifier, we used the R library *GBM*.

@@

The results for the two-class models in terms of loo-cross-validated accuracy and perplexity are:
\begin{table}
\begin{tabular}{|r|ccc|}
\hline
Model type & Loo-CV accuracy & Loo-CV perplexity&Binary accuracy on test data\\
\hline
Dummy model & 0 & 2.01&\\
Logistic regression & 0.8728448 & 1.338286&0.8445596\\
Naive Bayes & 0.8125 & 166.084 & 0.8 \\
LDA & 0.8922414 & 1.32923 &  0.8715026\\
QDA & 0.8534483 & 6.34633 & 0.8455959\\
SVM & 0.8793103 & -&0.8435233\\
Lasso & 0.8965517 & 1.29399&0.8746114\\
Random Forest & 0.887931 & 1.329754 & 0.8673575\\
GBM &	0.8943966 & 1.324385 & 0.8746114 \\
\hline
\end{tabular}
\end{table}

@@

\begin{table}[!htbp]
\begin{tabular}{|r|c|}
\hline
Explanatory variable & Estimate of coefficient\\
\hline
Intercept&16.16886\\
CO2168.std&0.02065910\\
CO242.mean&-0.04024650\\
CO2504.mean&-0.02066992\\
H2O42.mean&-0.2627714\\
H2O84.std&-2.407117e-01\\
NO168.std&2.281875\\
NOx672.std&-0.06064781\\
O3672.mean&0.09233133\\
O384.std&0.03595632\\
Pamb0.std&0.2960355\\
PTG.mean&-15.19914\\
RGlob.mean&0.001578964\\
RHIRGA168.std&0.01765280\\
RHIRGA336.mean&-0.02661884\\
RHIRGA672.mean&-0.03528919\\
RHIRGA672.std&0.06932797\\
SO2168.std&0.7092859\\
SWS.mean&0.01112000\\
T168.std&0.4947142\\
CS.mean&-938.8368\\
CS.std&580.6009\\
\hline
\end{tabular}
\end{table}

@@

## Deliverables

**For the deliverables, we give a grade 4.**
